\begin{center} \bf \sffamily \LARGE Prisoner's dilemma with stochastic payoffs\\[-1cm] \end{center}

### 2023-02-27 paper/main.tex
delete old parts
finish abstract

\begin{abstract}
Repeated games are vastly use in evolutionary game theory to explain cooperation
in a variety of environments. Although existing models have shaped our
understanding of human cooperation, hey often work with idealized assumptions.
In an evolutionary process, individuals imitate other individuals of the
population based on their fitness. It is commonly assumed that an individual
computes their fitness after interacting with a representative sample of the
population, and remembering all the interactions they participated in. In real
life, we do not always remember all our interactions, instead we rather recall
our most recent ones.

Here, we introduce a framework that allows individuals to estimate their fitness
based on a minimum of social information. We explore the difference between our
framework and the classical framework using computer simulations. The
simulations show that, in the prisoner's dilemma individuals with limited memory
tend to adopt less generous strategies and they achieve less cooperation than in
the classical scenario.
\end{abstract}

@@

\newpage
% In the
% simplest of cases each individual can be of one of two types, A and B. The state
% of the population is thus characterized by the number \(k\) of individuals of
% type A. The interaction between the two types of individuals is described by the
% functions \(\pi _{{\rm A}}^{i}\) and $\pi _{{\rm B}}^{i}$.

% airwise comparison process is a stochastic process for modelling the evolution
% of a finite population. The process starts with assigning all individuals of the
% population the same strategy. A strategy is a set of rules of how an individual
% should behave in an interaction with another individual. Each elementary time
% step of the process consists of three phases; (1) the \textbf{mutation phase}
% (2) the \textbf{game phase} and (3) the \textbf{update phase}. These are summarised
% in Figure~\ref{fig:pairwise_phases}.

% In the \textbf{mutation phase} one individual is
% chosen to switch to a new mutant strategy with a probability \(\mu\).
% In the \textbf{game phase} individuals are randomly matched with other
% individuals in the population, and they engage in a repeated game where each
% subsequent turn occurs with a fixed probability \(\delta\). At each turn the
% individuals decide on an action based on their strategies.
% In repeated games there are infinite many strategies, however, it is commonly
% assumed that individuals can only choose strategies from a restricted set. One
% such set is that of reactive strategies. A reactive
% strategy considers only the previous action of the other player, and thus, a
% reactive strategy \(s\) can be written as a three-dimensional vector \(s=(y, p,
% q)\). The parameter \(y\) is the probability that the strategy opens with a
% cooperation and \(p\), \(q\) are the probabilities that the strategy cooperates
% given that the opponent cooperated and defected equivalently.

% In the \textbf{update stage} where two individuals are randomly selected. From
% the two individuals, one  serves as the `learner' and the other as the `role
% model'. The learner adopts the role model's strategy with a probability \(\rho\)
% given by,

% The fitness of the individuals (\(\pi _{{\rm A}}^{i}, \pi _{{\rm B}}^{i}\)) is
% identified with the payoff resulting from the underline game. The evolution of
% cooperation remains one of the greatest problems for biological and social
% sciences. Cooperation is the action of choosing to help others at one's own
% expense. The standard game of formulating such situation is the Prisoner's
% Dilemma~\cite{trivers1971evolution, milinski1987tit, glynatsi2021bibliometric,
% ohtsuki2006simple}, in which two players can choose to cooperate or to defect.
% The players are offered a certain payoff, \(R\), for mutual cooperation and a
% lower payoff, \(P\), for mutual defection. If one player cooperates while the
% other defects, then the cooperator gets the lowest payoff, \(S\), while the
% defector gains the highest payoff, \(T\). Thus, the payoffs have the following
% property \(T > R > P > S\), making defection the dominant strategy in the
% non-repeated game.

% \begin{equation}\label{eq:payoffs}
%   \begin{blockarray}{ccc}
%       & \text{cooperate} & \text{defect} \\
%       \begin{block}{c(cc)}
%           \text{cooperate} & R & S \\
%           \text{defect} & T & P \\
%       \end{block}
%   \end{blockarray}
% \end{equation}

% In literature, the payoffs of A and B individuals only on the fraction of both
% types in the population. If there are \(k\) A individuals and \(N-k\) B
% individuals. In the context of the prisoner's dilemma the payoffs and assuming
% that A and B follow simple strategies where they cooperate and defect
% unconditionally respective, then the A and B individuals have fitness of

% \begin{equation}
%   \begin{array}{lcrcr}
%   \displaystyle \pi _{{\rm A}}^{i}	&=	&\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot R	&+	&\displaystyle\frac{k}{N-1}\cdot S,\\[0.5cm]
%   \displaystyle \pi _{{\rm B}}^{i}&=	&\displaystyle\frac{N-k}{N-1}\cdot T &+	&\displaystyle\frac{k-1}{N-1}\cdot P.\\
%   \end{array}
% \end{equation}

% respectively while self-interactions are excluded. In literature these are the
% expected payoffs. For \(\beta = 1\), fitness equals payoff. This scenario
% describes ``strong selection''. For \(\beta \ll 1\) , the payoff only provides a
% small perturbation to the overall fitness of an individual, a limit known as
% weak selection.

% In the evolutionary game theory literature it is usually assumed that players
% use strategies with finite memory~\cite{Nowak1992tit, Baek2016}. This assumption
% is common as it allows for an explicit calculation of the players'
% payoffs~\cite{sigmund2010calculus}. Players' payoffs are calculated in the limit
% of interactions assuming that the two strategies meet several times.

% Customarily, most evolutionary game studies make the additional assumption that
% individuals play many times and with all other players before reproduction takes
% place, so that payoffs, equivalently fitness, are given by the mean distribution
% of types in the population. This hypothesis, selection occurs much more slowly
% than the interaction between individuals and more that remember all the
% interactions they participated in. Although recent experimental studies show
% that this may not always be the case in biology [12– 14], it is clear that in
% cultural evolution or social learning the time scale of selection is much closer
% to the time scale of interaction. The effects of this mixing of scales cannot be
% disregarded [15], and then it is natural to ask about the consequences of the
% above assumption and the effect of relaxing it. The work of~\cite{Roca2006}.

% In this work we ask studies that make use of this classical framework report
% that cooperation can substantially evolve. However, the inconsistency in the
% memory size of players leads us to question the robustness of our understanding
% of cooperation. To this end, we propose a framework in which individuals,
% similar to the decisions at each turn, estimate their fitness based on a minimum
% of information.

% We first consider two extreme scenarios, the classical scenario and the
% alternative scenario where individuals update their strategies only based on the
% very last payoff they obtained. We observe that individuals with limited memory
% tend to adopt less generous strategies and they achieve less cooperation when
% interacting in a prisoner's dilemma. We obtain similar results when we consider
% that individuals update their strategies based on more information. More
% specifically, up to the last two payoffs they obtained when interacting with up
% to two different members of the population. We extend our approach to the rest
% of the symmetric \(2 \times 2\) games.

% The remainder of the paper is organized as follows. In
% section~\ref{section:model} we describe the model. In
% section~\ref{section:results} we present the results of the simulations, and in
% section~\ref{section:conclusions} we outline the main conclusions.

% In the following, we consider a well mixed population of fixed size. In each
% step of t

% We study the transmission of strategies with a frequency-dependent birth-death
% process26 in a finite population of size n. In each time step, two randomly
% chosen individuals compare their payoffs and one of them can switch to the other
% one's strategy. This process can be interpreted as a model for social learning,
% whereby successful strategies spread, and, occasionally, random strategy
% exploration introduces novel strategies (corresponding to mutations in
% biological models).

% We consider a population of \(N\) players~\footnote{The terms ``player'' and
% ``individual'' are used interchangeably here.} where \(N\) is even and mutations
% are sufficiently rare. Therefore, at any point in time there are at most two
% different strategies present in the population; a \textit{resident} strategy and
% a \textit{mutant} strategy. To describe how strategies spread we use a pairwise
% comparison process~\cite{Traulsen2006}. Each step of the evolutionary process
% consists of two stages, a game stage and an updating stage.

% In the game stage each individual is randomly matched with some other individual
% in the population. They engage in a match where each subsequent turn occurs with
% a fixed probability $\delta$. At each turn players choose independently to
% either cooperate (\(C\)) or to defect (\(D\)), and the payoffs of the turn
% depend on both their decisions. If both players cooperate they receive the
% reward payoff \(R\), whereas if both defect they receive the punishment payoff
% \(P\). If one cooperates but the other defects, the defector receives the
% temptation payoff \(T\), whereas the cooperator receives the sucker's payoff
% \(S\). We denote the feasible payoff of each turn as \(\mathcal{U} = \{R, S, T,
% P\}\). We assume that individuals use \textit{reactive strategies} to make
% decisions in each turn. Reactive strategies are a set of memory-one strategies
% that only take into account the previous action of the opponent. They can be
% written explicitly as a vector in \(\R_{3}\), more specifically, a reactive
% strategy \(s\) is given by \(s=(y, p, q)\). The parameter \(y\) is the
% probability that the strategy opens with a cooperation and \(p, q\) are the
% probabilities that the strategy cooperates given that the opponent cooperated
% and defected equivalently.

% In the updating stage, two players are randomly drawn from the population, a
% `learner' and a `role model'. Given the learner's payoff $u_L\!\in\!
% \mathcal{U}$ and the role model's payoff $u_{RL}\!\in\! \mathcal{U}$, the
% learner adopts the role model's strategy with probability,

% where $\beta\!\ge\!0$ is the strength of selection. For small values of $\beta$
% the imitation probability is independent of the strategies of the involved
% players. As the value of $\beta$ increases, the more likely it is that the
% learner adopts only strategies that yield a higher payoff. Conventionally the
% updating payoffs of the learner and the role model are based on their expected
% payoffs. A player's expected payoff is the mean payoff the player yields after
% engaging in matches of multiple turns with each member of the population. At
% each match a player bases their next turn decision only on the previous action
% of the opponent, however, the same player bases their expected payoffs on the
% outcomes of all their matches. Thus, a player is assumed to have limited and
% perfect memory at the same time. We propose a new a set of updating payoffs
% where it is also assumed that a player has limited memory. We referee to these
% as the limited memory payoffs.

% The evolutionary step is repeated until either the mutant strategy goes extinct,
% or until it fixes in the population. If the mutant fixes in the population then
% the mutant strategy becomes the new resident strategy. After either outcome we
% introduce a new mutant strategy uniformly chosen from all reactive strategies at
% random, and we set the number of mutants to $1$. This process of mutation and
% fixation/extinction is then iterated many times.

% \section{Results}\label{section:results} \subsection{Updating payoffs based on
% the last round with another member of the population}\label{section:donation}

% To investigate the role of extortion in the context of evolutionary games, we
% concentrate on the donation game.
% In this section we explore the case where the updating payoffs are based on the
% last round payoff achieved against another member of the population and we
% compare this to the classical scenario of the expected payoffs. We assume that
% each pair of players interacts in a donation game. The donation game is a
% special case of the prisoner's dilemma. Each player can choose to cooperate by
% providing a benefit \(b\) to the other player at their cost \(c\), with \(0 < c
% < b\). Thus, the feasible payoffs in each round are \(\mathcal{U} = \{b-c, -c,
% b, 0\}\).

% \begin{figure}[!htbp]
%   \centering
%   \includegraphics[width=\textwidth]{static/cooperation_rates_expected_and_stochastic_for_donation_game.pdf}
% \end{figure}

% We further explore the effect of benefit in
% Figure~\ref{fig:cooperation_rate_over_benefit}. The figure suggests that
% expected payoffs always yield a higher cooperation rate. In the case of expected
% payoffs we observe that the cooperation rate increases as the value of the
% benefit gets higher. In comparison for the limited memory payoffs, the
% cooperation rate remains unchanged at approximately 50\% once \(b=5\).

% % \begin{figure}[!htbp]
% %   \centering
% %   \begin{subfigure}{.5\textwidth}
% %     \centering
% %     \includegraphics[width=\textwidth]{static/expected_for_beta.pdf}
% %     \includegraphics[width=\textwidth]{static/stochastic_for_beta.pdf}
% %   \end{subfigure}%
% %   \begin{subfigure}{.5\textwidth}
% %     \centering
% %     \includegraphics[width=\textwidth]{static/cooperation_rate_over_b.pdf}
% %   \end{subfigure}
% %   \caption{{\bf The evolution of cooperation for different benefit values.} 
% %   We vary the benefit of defection $b$. In all cases, expected payoffs appear to
% %   overestimate the average cooperation rate the population achieves. ({\bf A})
% %   the probabilities \(p, q\) for resident population over \(10^7\) time steps
% %   for each benefit value. ({\bf B}) The cooperation rate within the resident
% %   population (averaged over all games and over all time steps) over the benefit.
% %   Unless explicitly varied, the parameters of the simulation are $N\!=\!100$,
% %   $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for
% %   $t\!=\!5\times 10^7$ time steps for each parameter
% %   combination.}\label{fig:cooperation_rate_over_benefit}
% % \end{figure}

% We also investigate the effect of the strength of selection $\beta$.
% Figure~\ref{fig:cooperation_rate_over_betas} illustrates results for various
% runs of the evolutionary process. For weak selection, \(\beta < 1\), we observe
% that the two methods yield similar results, however, as \(\beta\) increases
% there is variation in the evolving populations. In the case of expected payoffs
% the resident populations become more cooperative as \(\beta\) increases, whereas
% in the case of limited memory payoffs, the resident populations become more
% defective.

% % \begin{figure}[!htbp]
% %   \centering
% %     \includegraphics[width=\textwidth]{static/cooperation_rate_over_b.pdf}
% % \caption{{\bf The evolution of cooperation for different selection strength values.}
% % We vary the selection strength $\beta$. In all cases, stochastic payoff
% % evaluation tends to reduce the evolving cooperation rates. ({\bf A}) the
% % probabilities \(p, q\) for resident population over \(10^7\) time steps for each
% % \(\beta\) value. ({\bf B}) The cooperation rate within the resident population
% % (averaged over all games and over all time steps) over \(\beta\). Unless
% % explicitly varied, the parameters of the simulation are $N\!=\!100$, $b\!=\!3$,
% % $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for
% % $t\!=\!5\times 10^7$ time steps for each parameter
% % combination.}\label{fig:cooperation_rate_over_betas}
% % \end{figure}

% \begin{figure}[!htbp]
%   \centering
%     \includegraphics[width=\textwidth]{static/more_memory_summary_results}
% \caption{{\bf The evolution of cooperation for different selection strength values.}
% We vary the selection strength $\beta$. In all cases, stochastic payoff
% evaluation tends to reduce the evolving cooperation rates. ({\bf A}) the
% probabilities \(p, q\) for resident population over \(10^7\) time steps for each
% \(\beta\) value. ({\bf B}) The cooperation rate within the resident population
% (averaged over all games and over all time steps) over \(\beta\). Unless
% explicitly varied, the parameters of the simulation are $N\!=\!100$, $b\!=\!3$,
% $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for
% $t\!=\!5\times 10^7$ time steps for each parameter
% combination.}\label{fig:cooperation_rate_over_betas}
% \end{figure}

% \subsection{Effect of updating payoffs in different social
% dilemmas}\label{section:2_by_2_games}

% In the previous section we gained insights into the effects of the updating
% payoffs, and into how parameters such as the benefit and the strength of
% selection can intensify them. We investigated these effects by using the
% donation game. In order to broaden our understanding of the updating payoffs on
% different forms of possible human interactions we extend our approach to all \(2
% \times 2\) symmetric games. More specifically, we apply our analysis to four
% different classes of games; the harmony, the snowdrift, the prisoner's dilemma
% and the stag-hunt games.

% We compare the results of the evolutionary process when the updating payoffs are
% based on the expected payoffs, and on the last round payoffs for all the four
% possible classes of games (Figures~\ref{fig:expected_two_by_two}
% and~\ref{fig:last_round_two_by_two}). In Figure~\ref{fig:expected_two_by_two}
% each sub figure represents a run of evolutionary process for a different set of
% values for \(\mathcal{U}\). Without loss of generality we set \(R=1\) and
% \(P=0\)~\cite{Martinez2012, Roca2009}, and we vary the values of the temptation
% payoff \(T\) (across the \(x\)-axis) and of the sucker's payoff \(S\) (across
% the \(y\)-axis). Starting at the upper left corner and proceeding clockwise the
% quadrants correspond to the harmony, the snowdrift, the prisoner's dilemma and
% the stag-hunt games.

% % % \begin{figure}[!htbp]
% % %   \centering
% % %   \includegraphics[width=\textwidth]{static/expected_two_by_two_games.pdf}
% % %   \caption{{\bf Evolutionary dynamics under expected payoffs for various social dilemmas.}
% % %   We vary the temptation payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\) across
% % %   the \(x\) axis, and  \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\) across the
% % %   \(y\) axis. ({\bf A}) The top left quadrant where \(S > 0\) and \(T \leq 1\)
% % %   corresponds to the harmony game. The preference ordering for the harmony game
% % %   is \(R > T > S > P\). ({\bf B}) The top right quadrant where \(S > 0\) and \(T
% % %   > 1\) corresponds to the snowdrift game. The preference ordering for the
% % %   snowdrift game is \(T > R > S > P\). ({\bf C}) The bottom left quadrant where
% % %   \(S < 0\) and \(T \leq 1\) corresponds to the stag-hunt game. The preference
% % %   ordering for the stag-hunt game is \(R > T > P > S\). ({\bf D}) The bottom
% % %   right quadrant where \(S < 0\) and \(T > 1\) corresponds to the prisoner's
% % %   dilemma game. The preference ordering for the prisoner's dilemma game is \(T >
% % %   R > P > S\). Unless explicitly varied, the parameters of the simulation are
% % %   $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times 10^7$.}
% % %   \label{fig:expected_two_by_two}
% % % \end{figure}

% % % \begin{figure}[!htbp]
% % %   \centering
% % %   \includegraphics[width=\textwidth]{static/stochastic_two_by_two_games.pdf}
% % %   \caption{{\bf Evolutionary dynamics under last round payoffs for various social dilemmas.} 
% % %   We vary the temptation payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\) across
% % %   the \(x\) axis, and  \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\) across the
% % %   \(y\) axis. ({\bf A}) The top left quadrant where \(S > 0\) and \(T \leq 1\)
% % %   corresponds to the harmony game. The preference ordering for the harmony game
% % %   is \(R > T > S > P\). ({\bf B}) The top right quadrant where \(S > 0\) and \(T
% % %   > 1\) corresponds to the snowdrift game. The preference ordering for the
% % %   snowdrift game is \(T > R > S > P\). ({\bf C}) The bottom left quadrant where
% % %   \(S < 0\) and \(T \leq 1\) corresponds to the stag-hunt game. The preference
% % %   ordering for the stag-hunt game is \(R > T > P > S\). ({\bf D}) The bottom
% % %   right quadrant where \(S < 0\) and \(T > 1\) corresponds to the prisoner's
% % %   dilemma game. The preference ordering for the prisoner's dilemma game is \(T >
% % %   R > P > S\). Unless explicitly varied, the parameters of the simulation are
% % %   $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times 10^7$.}
% % %   \label{fig:last_round_two_by_two}
% % % \end{figure}

% % % The harmony game describes an interaction without conflict. In the harmony game
% % % it is in the best interest for both players to cooperate, and in an evolutionary
% % % process cooperators prevail. We observe that for most harmony game runs the
% % % resident population overwhelmingly applies a strategy for which \(p\) and \(q\)
% % % are \(\approx 1\) (Figure~\ref{fig:expected_two_by_two}). The snowdrift game
% % % describes a situation similar to that of the prisoner's dilemma; cooperation
% % % results in a benefit to the opposing player but entailing a cost to the
% % % cooperator. However, in the snowdrift game individuals obtain immediate direct
% % % benefits from the cooperative acts which leads to \(S > P\). In the snowdrift
% % % game more cooperation emerges compared to the prisoner's dilemma. Defection is
% % % never a resident strategy and for the same values of temptation the overall
% % % \(q-\) values are higher. The last class of games we present are for the
% % % stag-hunt game. In the stag-hunt game there is an equilibrium in which both
% % % players cooperate as well as one in which both defect. For small values of the
% % % temptation and the sucker's payoffs we observe that cooperation prevails similar
% % % to the harmony game. However, in the case where the temptation and the reward
% % % payoffs are equal the resident population either consists of defectors or
% % % cooperators.

% % % Although the patterns of the evolved populations remain similar if we consider
% % % that players update their strategies based on their last round payoff
% % % (Figure~\ref{fig:last_round_two_by_two}), there are still dissimilarities
% % % between the two approaches. To better understand these dissimilarities, we
% % % explore the cooperation rates for each of the cases of the social dilemmas we
% % % have presented (Figure~\ref{fig:cooperation_rates_two_by_two_part_one}).

% % % In the harmony game the cooperation rates for both approaches are similar. The
% % % biggest difference occurs for \(T=1\), nevertheless, it is less than 10\%. There
% % % are two instances for which the cooperation rates are higher for the limited
% % % memory payoffs compared to the expected. These are for \(T=2.2, S=0.4\) and for
% % % \(T=2.2, S=0.8\). It is only in the snowdrift game that cooperation is
% % % overestimated when a minimum of information is considered. Players are slightly
% % % more likely to cooperate after a cooperation and to forgive after being at the
% % % receiving end of a defection. For the rest of the cases the classical scenario
% % % overestimates the average cooperation rate, supporting the results of the
% % % previous section. There is a substantial difference in the cooperation rates for
% % % the prisoner's dilemma class. We observe that for \(S<-0.4\) cooperation almost
% % % never emerges if players imitate based only on the last round payoff.

% % % \begin{figure}[!htbp]
% % %   \centering
% % %   \includegraphics[width=.4\textwidth]{static/expected_two_by_two_games_cooperation.pdf}
% % %   \includegraphics[width=.4\textwidth]{static/stochastic_two_by_two_games_cooperation.pdf}
% % %   \caption{\textbf{Cooperation rates for various social dilemmas.} ({\bf A}) If
% % %   players update based on their expected payoffs. ({\bf B}) If players update
% % %   based on their last round payoffs. We vary the temptation payoff \(T \in
% % %   \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\) across the \(x\) axis, and \(S \in \{1.2,
% % %   0.8, 0.4, 0, -0.4, -0.8\}\) across the \(y\) axis. Unless explicitly varied,
% % %   the parameters of the simulation are $N\!=\!100$, $\beta\!=\!10$,
% % %   $\delta\!=\!0.99$, $t\!=\!5\times
% % %   10^7$.}\label{fig:cooperation_rates_two_by_two_part_one}
% % % \end{figure}

% % % So far we have explored the difference between the expected payoffs and the last
% % % round payoffs. In order to explore further the effect of limited memory we allow
% % % individuals to remember more. More precisely, up to two interactions and up to
% % % the last two rounds. In total we present results for three more updating
% % % payoffs. These are the payoffs when individuals consider the last two rounds
% % % with another member of the population, the last round with two members of the
% % % population, and the last two rounds with two members of the population. Similar
% % % to the last round payoff, we use simulations and record which strategies the
% % % players adopt over time based
% % % (Appendix~\ref{appendix:further_simulation_results}) and we compare the evolving
% % % cooperation rates to those of the expected payoffs
% % % (Figure~\ref{fig:cooperation_rates_two_by_two_part_two}).

% % % In the case of the last two rounds payoff we make two observations. Initially,
% % % for the harmony, the stag-hunt and the prisoner's dilemma classes the average
% % % cooperation rates are higher than the previous case of the last round payoff.
% % % The rates remain strictly less than in the expected payoffs, however, in the
% % % case of the prisoner's dilemma we observe a significant increase. Secondly, for
% % % the snowdrift class there are more instances for which the cooperation rates are
% % % higher than in the expected payoffs. This could indicate that in the case of the
% % % snowdrift game the expected payoffs underestimate cooperation. Once players
% % % interact with two other members of the population, instead of one, the expected
% % % payoffs strictly yield a higher cooperation rate  even in the case of the
% % % snowdrift game. Similar to the previous result, the difference between the two
% % % approaches is always higher for the prisoner's dilemma class.

% % % \begin{figure}[!htbp]
% % %   \centering
% % %   \includegraphics[width=.33\textwidth]{static/rounds_two_by_two_games_cooperation.pdf}
% % %   \includegraphics[width=.33\textwidth]{static/opponents_two_by_two_games_cooperation.pdf}
% % %   \includegraphics[width=.33\textwidth]{static/rounds_opponents_two_by_two_games_cooperation.pdf}
% % %   \caption{\textbf{Cooperation rates over various social dilemmas for different
% % %   limited memory payoffs.} ({\bf A}) If players update based on their last two
% % %   rounds with another member of the population. ({\bf B}) If players update
% % %   based on their last round with two other members of the population. ({\bf C})
% % %   If players update based on their last two rounds with two other members of the
% % %   population. We vary the temptation payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8,
% % %   2.2\}\) across the \(x\) axis, and \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\)
% % %   across the \(y\) axis. Unless explicitly varied, the parameters of the
% % %   simulation are $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times
% % %   10^7$..}
% % %   \label{fig:cooperation_rates_two_by_two_part_two}
% % % \end{figure}

% \section{Conclusions}\label{section:conclusions}

% ToDo: Add in discussion

% A major focus of evolutionary game theory is the evolution of cooperation among
% selfish individuals. Sometimes evolutionary game theory even seems to be reduced
% to the evolution of cooperation [Arne2022]. “Why do we act altruistically to
% provide a benefit to others when there is a cost to ourselves?” is still a
% puzzling question amongst researchers [Glynatsi]. In this determinist
% formulation, a small fraction of cooperative individuals that have a lower
% fitness can never invade a population of defecting strategies.  The stochastic
% effect can lead to fundamental changes since disadvantageous mutants have a
% small yet non-zero probability to reach fixation in a finite population, and
% thus cooperation can emerge.

% The framework of evolutionary game theory has been applied in explaining the
% behavior of sticklebacks~\cite{milinski1987tit}, the rock-scissors-paper cycles
% in bacterial populations~\cite{kerr2002local} and food sharing in vampire
% bats~\cite{carter2020development} colonies.

% Cooperation can be seen at odd, why is it that we choose to help others,
% increasing their payoff, at the expense of decreasing one's own? In spite of all
% the selfish genes’ animal and human communities seem to altruistically help each
% other and cooperate, and evolutionary game theory has helped us shape our
% understanding of the evolution of cooperation.

% Previous evolutionary models often feature a curious inconsistency. When
% modeling how individuals make decisions in each round, these models assume that
% players only remember the last round. However, when modeling how individuals
% update their strategies over time, individuals are assumed to have perfect
% memory.

% Here, we have explored how robust cooperation is as models deviate from the
% perfect memory assumption. Initially we considered the donation game. We showed
% that when the last round payoff is used instead of the expected payoffs,
% cooperation can even evolve when individuals only use a minimum of information,
% however, the evolving cooperation rates are typically lower. The resident
% strategies were both less cooperative and less generous. This effect was only
% intensified we increase the benefit and the strength of selection independently.
% The results showed that as each parameter was being increased the difference
% between the cooperation rates were widening. This indicates that cooperative
% players benefit from being able to interact with everyone in the population.

% We extended our approach and presented results not only based on different
% classes of social games, but also based on different limited memory payoffs. The
% analysis showed that specifically for the case of the prisoner's dilemma games
% the contrast between the two approaches can be significantly large. The
% prisoner's dilemma is one of the most applied types of social games when we
% discuss the evolution of cooperation. Our results indicate that cooperation
% struggles to evolve in the prisoner's dilemma when only a minimum of social
% information is used at the updating stage. Interestingly our simulations also
% showed that in same cases cooperation can benefit from limited memory. This was
% specific for the snowdrift game.

% \section{Acknowledgements}

% This work was supported by the European Research Council Starting Grant 850529:
% E-DIRECT.

>>>>

\begin{abstract}
Repeated games are vastly used in evolutionary game theory to explain cooperation
in a variety of environments. Although existing models have shaped our
understanding of human cooperation they often work with idealized assumptions.
In an evolutionary process, individuals imitate other individuals of the
population based on their fitness. It is commonly assumed that an individual
computes their fitness after interacting with a representative sample of the
population and remembering all the interactions they participated in. In real
life we do not always remember all our interactions, instead we rather recall
our most recent ones.

Here we introduce a framework that allows individuals to estimate their fitness
based on a minimum of social information. We explore the difference between our
framework and the classical framework using computer simulations. The
simulations show that in the prisoner's dilemma, individuals with limited memory
tend to adopt less generous strategies and they achieve less cooperation than in
the classical scenario.
\end{abstract}

@@

\newpage

\section{Conclusions}\label{section:conclusions}

A major focus of evolutionary game theory is the evolution of cooperation among
selfish individuals. Sometimes evolutionary game theory even seems to be reduced
to the evolution of cooperation [Arne2022]. “Why do we act altruistically to
provide a benefit to others when there is a cost to ourselves?” is still a
puzzling question amongst researchers [Glynatsi]. In this determinist
formulation, a small fraction of cooperative individuals that have a lower
fitness can never invade a population of defecting strategies.  The stochastic
effect can lead to fundamental changes since disadvantageous mutants have a
small yet non-zero probability to reach fixation in a finite population, and
thus cooperation can emerge.

Cooperation can be seen at odd, why is it that we choose to help others,
increasing their payoff, at the expense of decreasing one's own? In spite of all
the selfish genes' animal and human communities seem to altruistically help each
other and cooperate, and evolutionary game theory has helped us shape our
understanding of the evolution of cooperation.

Previous evolutionary models often feature a curious inconsistency. When
modeling how individuals make decisions in each round, these models assume that
players only remember the last round. However, when modeling how individuals
update their strategies over time, individuals are assumed to have perfect
memory.

Here, we have explored how robust cooperation is as models deviate from the
perfect memory assumption. Initially we considered the donation game. We showed
that when the last round payoff is used instead of the expected payoffs,
cooperation can even evolve when individuals only use a minimum of information,
however, the evolving cooperation rates are typically lower. The resident
strategies were both less cooperative and less generous. This effect was only
intensified we increase the benefit and the strength of selection independently.
The results showed that as each parameter was being increased the difference
between the cooperation rates were widening. This indicates that cooperative
players benefit from being able to interact with everyone in the population.

We extended our approach and presented results not only based on different
classes of social games, but also based on different limited memory payoffs. The
analysis showed that specifically for the case of the prisoner's dilemma games
the contrast between the two approaches can be significantly large. The
prisoner's dilemma is one of the most applied types of social games when we
discuss the evolution of cooperation. Our results indicate that cooperation
struggles to evolve in the prisoner's dilemma when only a minimum of social
information is used at the updating stage.

\section{Acknowledgements}

This work was supported by the European Research Council Starting Grant 850529:
E-DIRECT.
